FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-01 epss_marmo.tex
Update on Overleaf.

\begin{abstract}
Planetary science is a vast field of investigation that brings together several
research communities (geologists, astronomers, physicists, geochemists, etc.),
and produces an impressively growing amount of heterogeneous data.
Planetary surface research continues to evolve from mostly visual assessment
to more automated quantitative analysis.
Interoperability and openness of data formats and processing techniques are
becoming a necessity, to avoid the risk of being unable to efficiently extract
scientific information from the data, and to guarantee the reproducibility of
the scientific results.
This paper will describe how Flexible Image Transport System (FITS) can be used
in planetary surface investigations, and how its metadata can easily be inserted
in the PDS 4 metadata distribution model.
FITS format is open and flexible, largely implemented in open and efficient
processing tools.
Those reasons make FITS format appropriate for large amounts of raster data
processing.
The purpose of this approach is to make easier data mining and re-processing in
Planetary Surface investigations, promoting general software based on those
standards.
More than capturing a formal data model, FITS description for those data aims to
simplify operational logistics.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Planetary science is a vast field of investigation that brings together several
research communities (geologists, astronomers, physicists, geochemists, etc.),
and produces an impressively growing amount of heterogeneous data.
Planetary surface research continues to evolve from mostly visual assessment
to more automated quantitative analysis.
Interoperability and openness of data formats and processing techniques are
becoming a necessity, to avoid the risk of being unable to efficiently extract
scientific information from the data, and to guarantee the reproducibility of
the scientific results.
This paper will describe how Flexible Image Transport System (FITS) can be used
in planetary surface investigations, and how its metadata can easily be inserted
in the PDS 4 metadata distribution model.
FITS format is open and flexible, largely implemented in open and efficient
processing tools.
Those reasons make FITS format appropriate for large amounts of raster data
processing.
The purpose of this approach is to make easier data mining and re-processing in
Planetary Surface investigations, promoting general software based on those
standards.
More than capturing a formal data model, FITS description for those data aims to
simplify operational logistics.

The option to use FITS within the
planetary domain could be an opportunity to allow
more seamless sharing of data across these different
domains and potentially homogenize methods from
acquisition, to visualization, while giving more chances
to optimize data processing

This paper is organized as follows:
\begin{enumerate}
\item{a brief description of FITS format and on its pros and cons in using it in
planetary science;}
\item{a proposition for planetary surface metadata description in FITS}
\item{Dictionaries between FITS and PDS metadata description}
\item{examples of planetary dataset in FITS}
\item{current developments for making FITS usable in planetary sciences}
\end{enumerate}

@@

\subsection{FITS format}
FITS\footnote{https://fits.gsfc.nasa.gov/} \citep{fitsorig,fitsver3} stands for
'Flexible Image Transport System'.
It is an open digital standard, defined for data acquisition and archiving in astronomical
observatories back in the late 70's, and it is used for spatial telescope data.
In fact, FITS is the IAU-approved standard format for astronomical
data\footnote{https://fits.gsfc.nasa.gov/iaufwg/history/IAU\_1982\_resolution\_c1.html}. 
Therefore, FITS is one of the standard implemented formats in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html} (VO).
Finally, FITS format is compatible with the Planetary Data System (PDS)
3\footnote{https://pds.nasa.gov/pds4/about/} and
4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml} archiving
specifications and supported by a large number of open libraries and software tools.

>>>>

\begin{abstract}
Planetary science encompasses a broad number of research fields and
brings together several research communities (geologists, astronomers,
physicists, geochemists, etc.).
Planetary missions produce an impressively growing amount of heterogeneous
data which mining evolves those days from mostly visual assessment to more
automated quantitative analysis.
Interoperability and openness of data formats and processing techniques are
becoming a necessity, to avoid the risk of being unable to efficiently extract
scientific information from the data, and to guarantee the reproducibility of
the scientific results.
This paper will describe how Flexible Image Transport System (FITS) can be
used in planetary surface investigations, and how its metadata can easily be
inserted in the PDS4 metadata distribution model.
An evolution to the actual standard is proposed to adapt FITS to
planetary body description.
Some examples and implementations are then presented.
More than imposing a formal data model, FITS description for those data aims
to simplify the processing chain in planetary science, promoting
interoperability from data acquisition to final visualization.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
Planetary science brings together several research communities
(geologists, astronomers, physicists, geochemists, etc.),
and planetary missions produce a growing amount of heterogeneous data.
To efficiently extract scientific information from those data, they must
be made integrable, thus inter-operable.
Openness of data format and processing techniques is becoming a necessity,
to guarantee the reproducibility of the results obtained from those data.

This paper proposes to use Flexible Image Transport System\footnote{https://fits.gsfc.nasa.gov/}
(FITS) \citep{fitsorig,fitsver3} in planetary surface investigations,
and describes how FITS metadata can easily be inserted in the
PDS4\footnote{https://pds.nasa.gov/tools/standards-reference.shtml}
metadata distribution model.
FITS format is open and flexible, largely implemented in open and efficient
processing tools.
Those reasons make FITS format appropriate for large amounts of raster data
processing.
The goal of this approach is to make easier data mining and re-processing
in planetary surface investigations, promoting general software based on those
standards.
The option to use FITS within the planetary domain could potentially
homogenize methods from acquisition, to visualization, while
giving chances to optimize data processing.

This paper is organized as follows:
\begin{enumerate}
\item{a brief description of FITS format;}
\item{a proposition for planetary surface metadata description in FITS;}
\item{dictionaries between FITS and PDS metadata description;}
\item{some examples of planetary dataset in FITS;}
\item{current developments for making FITS usable in planetary sciences.}
\end{enumerate}

@@

\subsection{FITS format}
FITS stands for 'Flexible Image Transport System'.
It is an open digital standard, defined for data acquisition and archiving
in astronomical observatories back in the late 70's, and it is used for
spatial telescope data.
In fact, FITS is the IAU-approved standard format for astronomical
data\footnote{https://fits.gsfc.nasa.gov/iaufwg/history/IAU\_1982\_resolution\_c1.html}. 
Therefore, FITS is one of the standard implemented formats in the Virtual
Observatory\footnote{http://www.ivoa.net/astronomers/using\_the\_vo.html}
(VO).
Finally, FITS format is compatible with the Planetary Data System (PDS)
archiving specifications and supported by a large number of open libraries
and software tools.
